Rates and delay times of type Ia supernovae in the Dark Energy Survey

### 2021-05-20 draft.tex

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[P. Wiseman et al.]{
P. Wiseman,$^{1}$\thanks{E-mail: p.s.wiseman@soton.ac.uk}
M. Sullivan,$^{1}$
C. Frohmaier,$^{2,1}$
M. Smith, $^{3}$
M. Vincenzi, $^{2}$
et al.
\\
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

@@

% Abstract of the paper
\begin{abstract}
We use a sample of 809 photometrically classified type Ia supernovae (SNe Ia) discovered by the Dark Energy Survey (DES) along with 40415 field galaxies to calculate the rate of SNe Ia per galaxy in the redshift range $0.2 < z <0.6$. We recover the known correlation between SN Ia rate and galaxy stellar mass across a broad range of scales $8.5 \leq \log(M_*/\mathrm{M}_{\odot}) \leq 11.25$. We find that the SN Ia rate increases with stellar mass as a power-law with index $0.63 \pm 0.02$, which is consistent with previous work. We use an empirical model of stellar mass assembly to estimate the average star-formation histories (SFHs) of galaxies across the stellar mass range of our measurement. Combining the modelled SFHs with the SN Ia rates to estimate constraints on the SN Ia delay time distribution (DTD), we find the data are fit well by a power-law DTD with slope index $\beta = -1.14 \pm 0.05$ and normalisation $A = 2.1 \pm0.05 \times 10^{-13}~\mathrm{SNe}~{\mathrm{M}_{\odot}}^{-1}~\mathrm{yr}^{-1}$, which corresponds to an overall SN Ia production efficiency $N_{\mathrm{Ia}}/M_* = 0.9~_{-0.7}^{+4.0} \times 10^{-3}~\mathrm{SNe}~\mathrm{M}_{\odot}^{-1}$. Upon splitting the SN sample by properties of the light curves, we find a strong dependence on DTD slope with the SN decline rate, with slower-declining SNe exhibiting a steeper DTD slope. We interpret this as a result of a relationship between intrinsic luminosity and progenitor age, and explore the implications of the result in the context of SN Ia progenitors.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
keyword1 -- keyword2 -- keyword3
\end{keywords}

@@

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/mass_hist_linlog_sn_field_BC03.png}
    \caption{Upper: Distribution of the stellar mass of the SN host galaxies. The raw (green dashed) and spectroscopic-efficiency-weighted (blue dot-dashed; Section \ref{subsec:incompletenss_SN_hosts}) histograms are both normalised. Comparison samples are PTF (purple; \citealt{Pan2014}) and Pantheon (orange; \citealt{Scolnic2018}). Lower: As upper but for field galaxies, and shown on a log scale.  The distribution closely matches data from the ZFOURGE survey at $0.5<z<0.75$ (magenta dots). The difference between the raw counts (grey dotted) and the $V_{\mathrm{max}}$ corrected counts (orange solid; see Section \ref{sec:incompleteness}) is minimal. 
    \label{fig:mass_hist_hosts_field}}
\end{figure}

@@

    \item field galaxies must covered by $ugrizJHK$ photometry, with a photometric flux measurement or upper limit present in all bands;

@@

To correct for incompleteness caused by the magnitude limited nature of the survey we follow the prescription of e.g. \citet{Sullivan2006} and \citet{Smith2012} by using a $V_{\mathrm{max}}$ method based on \citet{Schmidt1968}. For each galaxy in the sample we calculate the maximum volume within which it would have been observed given its absolute magnitude and $k-$correction, and the apparent magnitude limits calculated using the $T_{\mathrm{C}}$ and $T_{\mathrm{V}}$ statistics. A correction of $V_{\mathrm{survey}}/V_{\mathrm{max}}$ is applied to all galaxies for which $V_{\mathrm{max}} < V_{\mathrm{survey}}$, where $V_{\mathrm{survey}}$ is the maximum volume reached by the survey. In our case this corresponds to the volume at $z=0.6$. Fig \ref{fig:vmax_field} shows the distribution of the correction among galaxies in the sample. The vast majority of galaxies require no correction, meaning they would have been observed beyond the maximum volume considered. Roughly 1\% of objects have a correction greater than 1, with the distribution well described by a power-law function. The effect of the $V_{\mathrm{max}}$ correction on the total galaxy counts in each stellar mass bin can be seen in Fig. \ref{fig:mass_hist_hosts_field}; indeed, the two histograms are visibly indistinguishable above $10^9~\mathrm{M}_{\odot}$ which indicates a sample that is effectively complete in this mass range in the redshift range of interest.

@@

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/SNuM.png}
    \caption{SN Ia rate per unit stellar mass as function of galaxy stellar mass. Error bars are the same as Fig. \ref{fig:rate_raw}. Literature data are that from SDSS spectra (orange squares; \citealt{Graur2013}) and the LOSS sample (grey dotted; \citealt{Li2011a}). The systematically higher SNuM in DES is caused by the higher redshift of the sample, as illustrated by the  prediction of the model using our best fit DTD from Section \ref{subsec:results_dtd} (coloured curves). }%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:snum}
\end{figure}
By binning both the SN hosts and field galaxies by their stellar mass we measure the mean SN Ia rate per galaxy as a function of stellar mass, $R_{\mathrm{G}}(M_*)$. We employ a bootstrap Monte-Carlo approach in order to estimate the uncertainty in the value of the rate in each stellar mass bin. The probability distribution function (PDF) of each galaxy's stellar mass is represented by the sum of two half Gaussian distributions to represent the asymmetric positive and negative uncertainties derived in the SED fitting stage (Section \ref{sec:data}). For each SN host and field galaxy, we take 100 samples of its stellar mass by drawing at random from its PDF. For each of the 100 samples we calculate the SN rate per galaxy in each stellar mass bin via Eq \ref{eq:rate1}, using the completeness-corrected values of $N_{\mathrm{SN}}$ and $N_{\mathrm{G}}$ from Eqs. \ref{eq:corr_SN} and \ref{eq:corr_G} respectively.

@@

A linear fit in log space with a best fitting slope of less than unity corresponds to a power-law, and indicates that SN Ia rate is not uniquely determined by host galaxy stellar mass. Indeed, the rate is instead likely to be driven by the DTD which is a non-linear function of stellar age. Investigating the shape of this DTD forms the basis for the latter part of this paper in Sections \ref{sec:model} and \ref{sec:split_x1_c}. The straight line fit in Fig. \ref{fig:rate_raw} deviates somewhat from the central values of the data at high stellar mass, the reason for which we investigate in Section \ref{subsec:discussion}.

@@

We begin by considering the delay time distribution of SNe Ia.
We represent the SN Ia DTD by a power law, which we consider effective after some "prompt time" $t_{\mathrm{p}}$, before which its value is set to 0. This is generally interpreted as the time taken for WDs to form after the burst of star formation. The DTD is thus:
\begin{equation}
 \Phi(\tau) = \left\{
    \begin{array}{@{}cc@{}}
    0 , & \tau < t_{\mathrm{p}} \\
    A\left(\frac{\tau}{\mathrm{Gyr}}\right)^{\beta} , & \tau \geq t_{\mathrm{p}}
    \end{array}\right.
        \label{eq:dtd}
\end{equation} 
where $\tau$ is the time since a burst of star formation, and $A$ is a normalisation at $\tau=1$ Gyr. If all stars in a galaxy were formed at a single epoch $t_f$, the DTD would describe the rate of SNe at an observation time $\tau = t_0$. However in reality galaxies are formed by the gradual build up of stellar mass over several epochs of star formation -- the distribution of this mass build up is known as the star-formation history (SFH). The rate of SNe Ia is thus the sum of the DTD evaluated for each epoch of star formation, multiplied by the stellar mass formed in that epoch. Mathematically this is represented by the convolution of the DTD and SFH:

@@

To model the SFH of galaxies as a function of their stellar mass we adopt the prescription of stellar mass assembly of \citetalias{Childress2014}, which draws on the work of \citet{Zahid2012}. In the model, galaxies are expected to evolve smoothly along the so-called "main sequence of star formation" whereby the SFR of a galaxy of stellar mass $M_*$ at redshift $z$ is determined by a simple relationship (the SMz relation, \citealt{Zahid2012}). Specifically, we implement their "fine-tuned" model (Eq. A7 in \citealt{Childress2014}), whereby the SMz flattens above $z\sim2$ in line with observations \citep{Stark2013}:

\begin{equation}
    \Psi(M_*,z) = \left(\frac{M_*}{10^{10}}\right)^{0.7}\frac{\exp{\left(1.9z\right)}}{\exp{\left(1.7\left(z-2\right)\right)} + \exp{\left(0.2\left(z-2\right)\right)} } [\mathrm{M}_{\odot} \mathrm{yr}^{-1}]\,,
    \label{eq:psi_mz}
\end{equation}
where $M_*$ is given in units of $\mathrm{M}_{\odot}$.
As galaxies grow, their SFR begins to slow down and eventually shut off almost entirely in a process known as quenching. We adopt the quenching penalty $p_Q$ directly from \citetalias{Childress2014}:
\begin{equation}
    p_Q(M_*,z) = \frac{1}{2}\left[1 - \mathrm{erf}\left(\frac{\log(M_*) - \log(M_Q(z))}{\sigma_Q}\right)\right]\,,
    \label{eq:pq}
\end{equation}
where $M_Q(z)$ describes how the quenching mass evolves with redshift, and $\sigma_Q$ is the transition scale which controls how fast a galaxy quenches. We adopt the observationally motivated form of the quenching mass evolution (Eq. A8 from \citealt{Childress2014}):
\begin{equation}
    \log(M_q(z)/\mathrm{M}_{\odot}) = 10.077 + 0.636z \,, 
    \label{eq:mq}
\end{equation}
and a transition scale of $\sigma_Q = 1.1$ which provides a good fit to data from the Galaxy And Mass Assembly (GAMA) survey \citep{Baldry2012}.

@@

As noted in \citetalias{Childress2014}, this toy model by construction follows several empirical relations derived from galaxy observations: model galaxies follow the SMz relation and obey quenching prescriptions motivated by observed galaxies. Here we make two further checks to validate to show that the model can be used to approximate the stellar age distribution of SN Ia host galaxies in DES.

@@

Secondly, we assess the shape of our SFHs and how they depend on stellar mass by comparing with SFH measurements from the Calar Alto Legacy Integral Field spectrscopy Area survey (CALIFA) as revealed by \citet{GonzalezDelgado2017}. Our models agree qualitatively on numerous features of the SFHs (their Fig. 5): high-mass galaxies form almost all their stars at high redshift; intermediate-mass galaxies display a plateau in their SFR from $\sim4$~Gyr to the present day; low-mass galaxies show an increasing rate of star formation at $z=0$. The main discrepancy with the lowest-mass bin of the CALIFA sample $8.6 <\log (M_*/\mathrm{M}_{\odot})<9.8$, which as well as an increasing SFR at low redshift shows a second peak of star formation at large lookback times, which we do not see in our model. However, this mass bin spans over one order of magnitude in stellar mass. Upon closer inspection of our model, the lower edge of this mass bin corresponds to galaxies that form at a lookback time of 5~Gyr and display a rising SFH. At the upper edge, galaxies do indeed form at times $>10$~Gyr and display an early peak followed by exponential decline. The combination of these different model SFHs in the low-mass bin thus explain the dual-peaked nature of the observed SFHs in CALIFA.

@@

SNe Ia do not occur instantaneously after an episode of star formation. Instead stars with zero-age main-sequence (ZAMS) masses less than $8~\mathrm{M}_{\odot}$ take time to evolve along the main sequence and form WDs and this time is not well constrained. In many works this ``prompt time" which we denote $t_{\mathrm{p}}$ is fixed to some value expected to be the minimum time for a star to evolve off the main sequence and become a WD, such as 40~Myr \citep{Maoz2012,Graur2013,Graur2014}. Other studies have left $t_{\mathrm{p}}$ as a free parameter \citep{Heringer2019,Castrillo2020}. We perform three fits: one with $t_{\mathrm{p}}$ fixed at 40 Myr, one with it as a free parameter along with $\beta$ and $A$, and a third with $\beta$ and $A$ fixed and  $t_{\mathrm{p}}$ free. We fit $t_{\mathrm{p}}$ in log space and adopt a Gaussian prior with hyper-parameters $p(\log(t_{\mathrm{p}})) \sim \mathcal{N}(-1.3\,,0.5)$, which corresponds to being centred around 50~Myr. Choosing this regime to fit allows more prior weight to be placed on shorter prompt times as per the majority of the literature. We use the $\hat{R}$ diagnostic of \citep{Vehtari2019} to assess the convergence of MC chains. When including $t_{\mathrm{p}}$ as a free parameter, we find that the strong degeneracy between $t_{\mathrm{p}}$ and $A$ cause the sampler to fail to converge with $\hat{R} \sim 1.3$ for the chains of those two parameters. In subsequent analyses we adopt as our fiducial model that with fixed $t_{\mathrm{p}} =0.04$~Gyr.

@@

The toy model of stellar mass assembly used in this work includes several assumptions about the evolution of galaxies. In particular, it is assumed that galaxies evolve independent of each other, growing up the star formation main sequence until they reach a mass at which they quench and star formation ceases. While these simplifications describe the average properties of galaxies well \citep{Zahid2012, Childress2014}, they struggle to describe the more stochastic nature of galaxy evolution that occurs at the low and high mass ends: starbursts and quenching episodes, respectively.

@@

In the previous sections we considered all SNe that passed the various quality, standardisation, and redshift cuts. However, it is well established that SNe Ia show correlations of varying strength between properties of their light curves and host galaxies. For example, it is well established that measures of the light curve duration (e.g. decline rate, or stretch such as SALT2 $x_1$) correlate with host galaxy properties such as morphology \citep{Hamuy1995,Hamuy2000,Mannucci2005}, stellar mass \citep{Kelly2010,Lampeitl2010,Sullivan2010} and specific SFR \citep{Rigault2013,Rigault2018}. If such correlations are caused by intrinsic differences in the progenitors or their surroundings, there may be signatures in the DTDs of SNe when divided into smaller samples. Such a hypothesis was investigated by \citet{Brandt2010} who found evidence for differing DTDs for SNe of low and high stretch, although \citet{Perrett2012} found no evidence for a change in volumetric rate evolution (expected due to the evolution of the CSFR) for the same stretch split. In this section we split the DES SNe Ia by their stretch and colour, and assess how the measured rates and DTDs vary across this parameter space. To do so we split the sample into sub-samples with $x_1 = 0$ and $c=0$ as the divisions points respectively. We repeat the analyses of Sections \ref{sec:rates} and \ref{sec:model} including all of the previous light curve cuts, performing the DTD fitting with a fixed $t_{\mathrm{p}} = 40$~Myr, and present the results in Table \ref{tab:dtd_results_split_lc}.

@@

All software used in this publication is publicly available. In particular, we made extensive use of \texttt{numpy} \citep{Harris2020}, \texttt{Astropy} \citep{AstropyCollaboration2013,AstropyCollaboration2018}, \texttt{matplotlib} \citep{Hunter2007}, \texttt{SciPy} \citep{Virtanen2020}, \texttt{pandas} \citep{Mckinney2010}, \texttt{Stan} \citep{Carpenter2017}, \texttt{seaborn} \citep{Waskom2020}, and \texttt{arviz} \citep{Kumar2019}.

\section*{Acknowledgements}

@@

The inclusion of a Data Availability Statement is a requirement for articles published in MNRAS. Data Availability Statements provide a standardised format for readers to understand the availability of data underlying the research results described in the article. The statement may refer to original data generated in the course of the study or to third-party data analysed in the article. The statement should describe and provide means of access, where possible, by linking to the data or providing the required accession numbers for the relevant databases or DOIs.

>>>>

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[P. Wiseman et al.]{\parbox{\textwidth}{
\Large
P.~Wiseman,$^{1}$
M.~Sullivan,$^{1}$
M.~Smith,$^{2,1}$
C.~Frohmaier,$^{1,3}$
M.~Vincenzi,$^{3}$
O.~Graur,$^{3}$
B.~Popovic,$^{4}$
P.~Armstrong,$^{5}$
D.~Brout,$^{6,7}$
D.~Carollo,$^{8}$
T.~M.~Davis,$^{9}$
L.~Galbany,$^{10}$
S.~R.~Hinton,$^{9}$
L.~Kelsey,$^{1}$
R.~Kessler,$^{11,12}$
G.~F.~Lewis,$^{13}$
C.~Lidman,$^{14,5}$
A.~M\"oller,$^{15}$
R.~C.~Nichol,$^{3}$
B.~Rose,$^{4}$
D.~Scolnic,$^{4}$
M.~Toy,$^{1}$
B.~E.~Tucker,$^{5}$
Z.~Zontou,$^{1}$
J.~Asorey,$^{16}$
T.~M.~C.~Abbott,$^{17}$
M.~Aguena,$^{18,19}$
S.~Allam,$^{20}$
F.~Andrade-Oliveira,$^{21,19}$
J.~Annis,$^{20}$
D.~Bacon,$^{3}$
E.~Bertin,$^{22,23}$
D.~Brooks,$^{24}$
E.~Buckley-Geer,$^{11,20}$
D.~L.~Burke,$^{25,26}$
A.~Carnero~Rosell,$^{27,19,28}$
M.~Carrasco~Kind,$^{29,30}$
J.~Carretero,$^{31}$
M.~Costanzi,$^{32,33,34}$
L.~N.~da Costa,$^{19,35}$
M.~E.~S.~Pereira,$^{36}$
S.~Desai,$^{37}$
H.~T.~Diehl,$^{20}$
P.~Doel,$^{24}$
S.~Everett,$^{38}$
I.~Ferrero,$^{39}$
B.~Flaugher,$^{20}$
P.~Fosalba,$^{40,41}$
J.~Frieman,$^{20,12}$
J.~Garc\'ia-Bellido,$^{42}$
E.~Gaztanaga,$^{40,41}$
T.~Giannantonio,$^{43,44}$
D.~Gruen,$^{45,25,26}$
R.~A.~Gruendl,$^{29,30}$
J.~Gschwend,$^{19,35}$
G.~Gutierrez,$^{20}$
D.~L.~Hollowood,$^{38}$
K.~Honscheid,$^{46,47}$
B.~Hoyle,$^{48,49}$
D.~J.~James,$^{7}$
E.~Krause,$^{50}$
K.~Kuehn,$^{51,52}$
N.~Kuropatkin,$^{20}$
M.~A.~G.~Maia,$^{19,35}$
J.~L.~Marshall,$^{53}$
P.~Martini,$^{46,54,55}$
F.~Menanteau,$^{29,30}$
R.~Miquel,$^{56,31}$
R.~Morgan,$^{57}$
R.~L.~C.~Ogando,$^{19,35}$
A.~Palmese,$^{20,12}$
F.~Paz-Chinch\'{o}n,$^{29,43}$
D.~Petravick,$^{29}$
A.~Pieres,$^{19,35}$
A.~A.~Plazas~Malag\'on,$^{58}$
A.~K.~Romer,$^{59}$
E.~Sanchez,$^{16}$
V.~Scarpine,$^{20}$
M.~Schubnell,$^{36}$
S.~Serrano,$^{40,41}$
I.~Sevilla-Noarbe,$^{16}$
E.~Suchyta,$^{60}$
M.~E.~C.~Swanson,$^{29}$
G.~Tarle,$^{36}$
D.~Thomas,$^{3}$
C.~To,$^{45,25,26}$
T.~N.~Varga,$^{49,61}$
and A.~R.~Walker$^{17}$
\begin{center} (DES Collaboration) \end{center}
}
\vspace{0.4cm}
\\
\parbox{\textwidth}{Affiliations are listed at the end of the paper}
}
% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

@@

% Abstract of the paper
\begin{abstract}
We use a sample of 809 photometrically classified type Ia supernovae (SNe Ia) discovered by the Dark Energy Survey (DES) along with 40415 field galaxies to calculate the rate of SNe Ia per galaxy in the redshift range $0.2 < z <0.6$. We recover the known correlation between SN Ia rate and galaxy stellar mass across a broad range of scales $8.5 \leq \log(M_*/\mathrm{M}_{\odot}) \leq 11.25$. We find that the SN Ia rate increases with stellar mass as a power-law with index $0.63 \pm 0.02$, which is consistent with previous work. We use an empirical model of stellar mass assembly to estimate the average star-formation histories (SFHs) of galaxies across the stellar mass range of our measurement. Combining the modelled SFHs with the SN Ia rates to estimate constraints on the SN Ia delay time distribution (DTD), we find the data are fit well by a power-law DTD with slope index $\beta = -1.13 \pm 0.05$ and normalisation $A = 2.11 \pm0.05 \times 10^{-13}~\mathrm{SNe}~{\mathrm{M}_{\odot}}^{-1}~\mathrm{yr}^{-1}$, which corresponds to an overall SN Ia production efficiency $N_{\mathrm{Ia}}/M_* = 0.9~_{-0.7}^{+4.0} \times 10^{-3}~\mathrm{SNe}~\mathrm{M}_{\odot}^{-1}$. Upon splitting the SN sample by properties of the light curves, we find a strong dependence on DTD slope with the SN decline rate, with slower-declining SNe exhibiting a steeper DTD slope. We interpret this as a result of a relationship between intrinsic luminosity and progenitor age, and explore the implications of the result in the context of SN Ia progenitors.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
supernovae: general -- galaxies: evolution -- white dwarfs
\end{keywords}

@@

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/mass_hist_linlog_sn_field_BC03.png}
    \caption{Upper: Distribution of the stellar mass of the SN host galaxies. The raw (green dashed) and spectroscopic-efficiency-weighted (blue dot-dashed; Section \ref{subsec:incompletenss_SN_hosts}) histograms are both normalised. Comparison samples are PTF (purple; \citealt{Pan2014}) and Pantheon (orange; \citealt{Scolnic2018}). Lower: As upper but for field galaxies, and with galaxy frequency shown on a log scale.  The distribution closely matches data from the ZFOURGE survey at $0.5<z<0.75$ (magenta dots). The difference between the raw counts (grey dotted) and the $V_{\mathrm{max}}$ corrected counts (orange solid; see Section \ref{sec:incompleteness}) is minimal. 
    \label{fig:mass_hist_hosts_field}}
\end{figure}

@@

    \item field galaxies must be covered by $ugrizJHK$ photometry, with a photometric flux measurement or upper limit present in all bands;

@@

To correct for incompleteness caused by the magnitude limited nature of the survey we follow the prescription of \citet{Sullivan2006} and \citet{Smith2012} by using a $V_{\mathrm{max}}$ method based on \citet{Schmidt1968}. For each galaxy in the sample we calculate the maximum volume within which it would have been observed given its absolute magnitude and $k-$correction, and the apparent magnitude limits calculated using the $T_{\mathrm{C}}$ and $T_{\mathrm{V}}$ statistics. A correction of $V_{\mathrm{survey}}/V_{\mathrm{max}}$ is applied to all galaxies for which $V_{\mathrm{max}} < V_{\mathrm{survey}}$, where $V_{\mathrm{survey}}$ is the maximum volume reached by the survey. In our case this corresponds to the volume at $z=0.6$. Fig \ref{fig:vmax_field} shows the distribution of the correction among galaxies in the sample. The vast majority of galaxies require no correction, meaning they would have been observed beyond the maximum volume considered. Roughly 1\% of objects have a correction greater than 1, with the distribution well described by a power-law function. The effect of the $V_{\mathrm{max}}$ correction on the total galaxy counts in each stellar mass bin can be seen in Fig. \ref{fig:mass_hist_hosts_field}; indeed, the two histograms are visibly indistinguishable above $10^9~\mathrm{M}_{\odot}$ which indicates a sample that is effectively complete in this mass range in the redshift range of interest.

@@

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/SNuM.png}
    \caption{SN Ia rate per unit stellar mass as function of galaxy stellar mass. Error bars are the same as Fig. \ref{fig:rate_raw}. Literature data are that from SDSS spectra (orange squares; \citealt{Graur2013}) and the LOSS sample (grey dotted; \citealt{Li2011a}). The systematically higher SNuM in DES is caused by the higher redshift of the sample, as illustrated by the  prediction of the model using our best fit DTD from Section \ref{subsec:results_dtd} (coloured curves). }%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:snum}
\end{figure}
By binning both the SN hosts and field galaxies by their stellar mass we measure the mean SN Ia rate per galaxy as a function of stellar mass, $R_{\mathrm{G}}(M_*)$. We employ a bootstrap Monte-Carlo approach in order to estimate the uncertainty in the value of the rate in each stellar mass bin. The probability density function (PDF) of each galaxy's stellar mass is represented by the sum of two half Gaussian distributions to represent the asymmetric positive and negative uncertainties derived in the SED fitting stage (Section \ref{sec:data}). For each SN host and field galaxy, we take 100 samples of its stellar mass by drawing at random from its PDF. For each of the 100 samples we calculate the SN rate per galaxy in each stellar mass bin via Eq \ref{eq:rate1}, using the completeness-corrected values of $N_{\mathrm{SN}}$ and $N_{\mathrm{G}}$ from Eqs. \ref{eq:corr_SN} and \ref{eq:corr_G} respectively.

@@

A best fitting slope of less than unity corresponds to a power-law, and indicates that SN Ia rate is not uniquely determined by host galaxy stellar mass. Indeed, the rate is instead likely to be driven by the DTD which is a non-linear function of stellar age. Investigating the shape of this DTD forms the basis for the latter part of this paper in Sections \ref{sec:model} and \ref{sec:split_x1_c}. The straight line fit in Fig. \ref{fig:rate_raw} deviates somewhat from the central values of the data at high stellar mass, the reason for which we investigate in Section \ref{subsec:discussion}.

@@

We begin by considering the delay time distribution of SNe Ia.
We represent the SN Ia DTD by a power law, which we consider effective after some "prompt time" $t_{\mathrm{p}}$, before which its value is set to 0. $t_{\mathrm{p}}$ is generally interpreted as the time taken for WDs to form after the burst of star formation. The DTD is thus:
\begin{equation}
 \Phi(\tau) = \left\{
    \begin{array}{@{}cc@{}}
    0 , & \tau < t_{\mathrm{p}} \\
    A\left(\frac{\tau}{\mathrm{Gyr}}\right)^{\beta} , & \tau \geq t_{\mathrm{p}}
    \end{array}\right.
        \label{eq:dtd}
\end{equation} 
where $\tau$ is the time since a burst of star formation, and $A$ is a normalisation at $\tau=1$ Gyr. If all stars in a galaxy were formed at a single epoch $t_f$, the DTD would describe the rate of SNe at an observation time $\tau = t_0$. However in reality galaxies are formed by the gradual build up of stellar mass over several epochs of star formation -- the distribution of this mass build up is known as the star-formation history (SFH). The rate of SNe Ia is thus the sum of the DTD evaluated for each epoch of star formation, multiplied by the stellar mass formed in that epoch. Mathematically this is represented by the convolution of the DTD and SFH:

@@

To model the SFH of galaxies as a function of their stellar mass we adopt the prescription of stellar mass assembly of \citetalias{Childress2014}, which draws on the work of \citet{Zahid2012}. In the model, galaxies are expected to evolve smoothly along the so-called "main sequence of star formation" whereby the SFR of a galaxy of stellar mass $M_*$ at redshift $z$ is determined by a simple relationship (the SMz relation, \citealt{Zahid2012}). Specifically, we implement their "fine-tuned" model (Eq. A7 in \citetalias{Childress2014}), whereby the SMz flattens above $z\sim2$ in line with observations \citep{Stark2013}:

\begin{equation}
    \Psi(M_*,z) = \left(\frac{M_*}{10^{10}}\right)^{0.7}\frac{\exp{\left(1.9z\right)}}{\exp{\left(1.7\left(z-2\right)\right)} + \exp{\left(0.2\left(z-2\right)\right)} } [\mathrm{M}_{\odot} \mathrm{yr}^{-1}]\,,
    \label{eq:psi_mz}
\end{equation}
where $M_*$ is given in units of $\mathrm{M}_{\odot}$.
As galaxies grow, their SFR begins to slow down and eventually shut off almost entirely in a process known as quenching. We adopt the quenching penalty $p_Q$ directly from \citetalias{Childress2014}:
\begin{equation}
    p_Q(M_*,z) = \frac{1}{2}\left[1 - \mathrm{erf}\left(\frac{\log(M_*) - \log(M_Q(z))}{\sigma_Q}\right)\right]\,,
    \label{eq:pq}
\end{equation}
where $M_Q(z)$ describes how the quenching mass evolves with redshift, and $\sigma_Q$ is the transition scale which controls how fast a galaxy quenches. We adopt the observationally motivated form of the quenching mass evolution (Eq. A8 from \citetalias{Childress2014}):
\begin{equation}
    \log(M_q(z)/\mathrm{M}_{\odot}) = 10.077 + 0.636z \,, 
    \label{eq:mq}
\end{equation}
and a transition scale of $\sigma_Q = 1.1$ which provides a good fit to data from the Galaxy And Mass Assembly (GAMA) survey \citep{Baldry2012}.

@@

As noted in \citetalias{Childress2014}, this toy model by construction follows several empirical relations derived from galaxy observations: model galaxies follow the SMz relation and obey quenching prescriptions motivated by observed galaxies. Here we make two further checks to validate that the model can be used to approximate the stellar age distribution of SN Ia host galaxies in DES.

@@

Secondly, we assess the shape of our SFHs and how they depend on stellar mass by comparing with SFH measurements from the Calar Alto Legacy Integral Field spectroscopy Area survey (CALIFA) as revealed by \citet{GonzalezDelgado2017}. Our models agree qualitatively on numerous features of the SFHs (their Fig. 5): high-mass galaxies form almost all their stars at high redshift; intermediate-mass galaxies display a plateau in their SFR from $\sim4$~Gyr to the present day; low-mass galaxies show an increasing rate of star formation at $z=0$. The main discrepancy with the lowest-mass bin of the CALIFA sample $8.6 <\log (M_*/\mathrm{M}_{\odot})<9.8$, which as well as an increasing SFR at low redshift shows a second peak of star formation at large lookback times, which we do not see in our model. However, this mass bin spans over one order of magnitude in stellar mass. Upon closer inspection of our model, the lower edge of this mass bin corresponds to galaxies that form at a lookback time of 5~Gyr and display a rising SFH. At the upper edge, galaxies do indeed form at times $>10$~Gyr and display an early peak followed by exponential decline. The combination of these different model SFHs in the low-mass bin thus explain the dual-peaked nature of the observed SFHs in CALIFA.

@@

SNe Ia do not occur instantaneously after an episode of star formation. Instead stars with zero-age main-sequence (ZAMS) masses less than $8~\mathrm{M}_{\odot}$ take time to evolve along the main sequence and form WDs and this time is not well constrained. In many works this ``prompt time" which we denote $t_{\mathrm{p}}$ is fixed to some value expected to be the minimum time for a star to evolve off the main sequence and become a WD, such as 40~Myr \citep{Maoz2012,Graur2013,Graur2014}. Other studies have left $t_{\mathrm{p}}$ as a free parameter \citep{Heringer2019,Castrillo2020}. We perform three fits: one with $t_{\mathrm{p}}$ fixed at 40 Myr, one with it as a free parameter along with $\beta$ and $A$, and a third with $\beta$ and $A$ fixed and  $t_{\mathrm{p}}$ free. We fit $t_{\mathrm{p}}$ in log space and adopt a Gaussian prior with hyper-parameters $p(\log(t_{\mathrm{p}})) \sim \mathcal{N}(-1.3\,,0.5)$, which corresponds to being centred around 50~Myr. Choosing this regime to fit allows more prior weight to be placed on shorter prompt times as per the majority of the literature. We use the $\hat{R}$ diagnostic of \citet{Vehtari2019} to assess the convergence of MC chains. When including $t_{\mathrm{p}}$ as a free parameter, we find that the strong degeneracy between $t_{\mathrm{p}}$ and $A$ cause the sampler to fail to converge with $\hat{R} \sim 1.3$ for the chains of those two parameters. In subsequent analyses we adopt as our fiducial model that with fixed $t_{\mathrm{p}} =0.04$~Gyr.

@@

The toy model of stellar mass assembly used in this work includes several assumptions about the evolution of galaxies. In particular, it is assumed that galaxies evolve independent of each other, growing up the star formation main sequence until they reach a mass at which they quench and star formation ceases. While these simplifications describe the average properties of galaxies well (\citealt{Zahid2012},\citetalias{Childress2014}), they struggle to describe the more stochastic nature of galaxy evolution that occurs at the low and high mass ends: starbursts and quenching episodes, respectively.

@@

In the previous sections we considered all SNe that passed the various quality, standardisation, and redshift cuts. However, it is well established that SNe Ia show correlations of varying strength between properties of their light curves and host galaxies. For example, it is well established that measures of the light curve duration (e.g. decline rate, or stretch such as SALT2 $x_1$) correlate with host galaxy properties such as morphology \citep{Hamuy1995,Hamuy2000,Mannucci2005}, stellar mass \citep{Kelly2010,Lampeitl2010,Sullivan2010} and specific SFR \citep{Rigault2013,Rigault2018}. If such correlations are caused by intrinsic differences in the progenitors or their surroundings, there may be signatures in the DTDs of SNe when divided into smaller samples. Such a hypothesis was investigated by \citet{Brandt2010} who found evidence for differing DTDs for SNe of low and high stretch, although \citet{Perrett2012} found no evidence for a change in volumetric rate evolution (expected due to the evolution of the CSFR) for the same stretch split. In this section we split the DES SNe Ia by their stretch and colour, and assess how the measured rates and DTDs vary across this parameter space. To do so we split the sample into sub-samples with $x_1 = 0$ and $c=0$ as the divisions points respectively. We repeat the analyses of Sections \ref{sec:rates} and \ref{sec:model} including all of the previous light curve cuts, performing the DTD fitting with a fixed $t_{\mathrm{p}} = 40$~Myr, and present the results in Table \ref{tab:dtd_results_split_lc}. Posterior distributions for fitted parameters are presented in Appendix \ref{appendix:posteriors}.

@@

All software used in this publication is publicly available. In particular, we made extensive use of \texttt{numpy} \citep{Harris2020}, \texttt{Astropy} \citep{AstropyCollaboration2013,AstropyCollaboration2018}, \texttt{matplotlib} \citep{Hunter2007}, \texttt{SciPy} \citep{Virtanen2020}, \texttt{pandas} \citep{Mckinney2010}, \texttt{Stan} \citep{Carpenter2017}, \texttt{seaborn} \citep{Waskom2020}, and \texttt{ChainConsumer} \citep{Hinton2016}.

\section*{Acknowledgements}
We thank J. Michael Burgess for discussions regarding model fitting in Stan.

 P.W. acknowledges support from the Science and Technology Facilities Council (STFC) grant ST/R000506/1. P.W., M.Sullivan and M.Smith acknowledge support from EU/FP7-ERC grant 615929.

@@

The DES-SN photometric SN Ia catalogue and associated host galaxy data will be made available as part of the DES5YR SN cosmology analysis at https://des.ncsa.illinois.edu/releases/sn. Field galaxy data including photometric redshifts will appear alongside other data products released as part of the DES Y3 weak lensing cosmology analysis at https://des.ncsa.illinois.edu/releases.
